---
title: "Homework-8"
author: Bryan Sandor, Merilyn Thomas, and Tyler White
format: html
editor: visual
execute:
  warning: false
  message: false
---

## From HW 7

```{r}
library(tidyverse)
library(readr)
library(lubridate)
library(janitor)
library(corrplot)
library(tidymodels)
```

## Basic Modeling Practice

### Reading Data

In this homework, we will do some structured practice with fitting linear models in R. The dataset from the UCI Machine Learning Repository containing daily bike sharing rentals in Seoul. Our goal is to perform an exploratory data analysis, clean the data, and fit various multiple linear regression models to predict daily bike counts.

```{r}
bike_data <- read_csv("SeoulBikeData.csv", locale = locale(encoding="latin1"))
```

### EDA

#### Data Processing

Before fitting any models, conduct basic data cleaning to prepare our data for usage. First, check for missing values and verify data types to make sure the columns were read in correctly. Check the column types and the values within the columns go make sure they make sense.

```{r}
# Check structure of data
str(bike_data)

# Check for missing values
colSums(is.na(bike_data)) 
```

None of the columns have missing values! Note that the `Date` column has been read in as a character type, so this column needs to be processed further. The categorical variables `Seasons`, `Holiday`, and `Functioning Day` have also been read in as character type, so these will be converted to factor type for later analysis. The `Date` column is currently stored as a character string with the format day/month/year. This will be converted into a usable date object using the `lubridate` library.

```{r Date Conversion}
#Convert the Date column into an actual date 
bike_clean <- bike_data |>
  mutate(Date = dmy(Date))
```

Turn the categorical variables and make sure they are read as factors.

```{r Characters to Factors}
#Turn the character variables Seasons, Holiday, and Functioning Day) into factors.
bike_clean <- bike_clean |>
  mutate(Seasons = as.factor(Seasons),
         Holiday = as.factor(Holiday),
         `Functioning Day` = as.factor(`Functioning Day`))

head(bike_clean)
```

Rename all the variables to easy names.

```{r}
bike_clean <- bike_clean |>
  clean_names() # Replaces all spaces with '_' and makes all letters lowercase.
names(bike_clean)
```

#### Exploratory Data Analysis

Start completing basic exploratory data analysis (EDA). Create summary statistics as they relate to the bike rental count. View these statistics across some categorical variables as well.

```{r}
# Summary of rented_bike_count variable.
summary(bike_clean$rented_bike_count)

# Find unique values for each categorical variable
unique(bike_clean$seasons)
unique(bike_clean$holiday)
unique(bike_clean$functioning_day)
```

Since `functioning_day` variable only has Yes or No values, this means that the bike sharing system was either operable or inoperable that hour. If the system was inoperable, then there is no useful information about bike rentals for that hour. In this case, remove observations that show the system was not operating.

```{r}
table(bike_clean$functioning_day) # 295 "No " observations

bike_clean <- bike_clean |>
  filter(functioning_day == "Yes")
```

The data is currently given as hourly observations. Summarize across the hours so that each day has one observation associated with it. Group the data by `date`, `seasons`, and `holiday`. Find the sum of the `rented_bike_count`, `rainfall`, and `snowfall` variables. Include the mean of all the weather related variables.

```{r}
daily_bike_data <- bike_clean %>%
  group_by(date, seasons, holiday) %>%
  summarise(
    total_bike_count = sum(rented_bike_count),
    avg_temp_c = mean(temperature_c, na.rm = TRUE),
    avg_humidity_pct = mean(humidity_percent, na.rm = TRUE),
    avg_windspeed_m_s = mean(wind_speed_m_s, na.rm = TRUE),
    avg_visibility_10m = mean(visibility_10m, na.rm = TRUE),
    avg_dew_point_temp_c = mean(dew_point_temperature_c, na.rm = TRUE),
    avg_solar_radiation_mj_m2 = mean(solar_radiation_mj_m2, na.rm = TRUE),
    total_rainfall_mm = sum(rainfall_mm, na.rm = TRUE),
    total_snowfall_cm = sum(snowfall_cm, na.rm = TRUE)
  )

head(daily_bike_data)
```

The `daily_bike_data` tibble will be the new data that will be analyzed! Recreate the basic summary statistics and then create some plots to explore variable relationships. View a correlation matrix between numeric variables to see if the data shows any significant correlations.

```{r}
# Summary of total_bike_count variable on a given day of the year.
summary(daily_bike_data$total_bike_count)
```

#### Graphs

Now let's examine some plots. First, let's relate rainfall to total rentals.

```{r rainfall_v_rentals_graph}
ggplot(data = daily_bike_data,
       aes(x = total_rainfall_mm, y = total_bike_count)) +
  geom_jitter(width = 1) +
  geom_smooth(method = "lm") +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 40000)) +
  labs(x = "Total Rainfall (mm)",
       y = "Total Bike Rentals",
       title = "Rainfall vs Bike Rentals")
```

It looks like an inverse relation exists between rainfall and rentals. Now let's do the same but for snowfall. One would assume a similar relationship exists.

```{r}
ggplot(data = daily_bike_data,
       aes(x = total_snowfall_cm, y = total_bike_count)) +
  geom_jitter(width = 1) +
  geom_smooth(method = "lm") +
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 40000)) +
  labs(x = "Total Snowfall (cm)",
       y = "Total Bike Rentals",
       title = "Snowfall vs Bike Rentals")
```

Not surprisingly, this seems more stark; after all, snow is more dangerous for bike rides and some people even prefer to ride in the rain. Let's examine the rentals across each of the four seasons.

```{r}
ggplot(data = daily_bike_data,
       aes(x = seasons, y = total_bike_count, fill = seasons)) +
  geom_col() +
  labs(x = "Season",
       y = "Total Bike Rentals",
       title = "Seasonal Bike Rentals") +
  theme(legend.position = "none")
```

It appears people prefer sunny weather or cooler, but not frigid, temperatures, and this is practical. Let's touch on this last thought with two more comparisons: solar radiation and temperature, respectively.

```{r}
ggplot(data = daily_bike_data,
       aes(x = avg_solar_radiation_mj_m2, y = total_bike_count)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm") +
  labs(x = "Solar Radiation (MJ/m^2)",
       y = "Total Bike Rentals",
       title = "Solar Radiation vs Bike Rentals")
```

Indeed, sunnier weather seems directly proportional to bike rentals, albeit not within as narrow a band as our rain or snow fall.

```{r}
ggplot(data = daily_bike_data,
       aes(x = avg_temp_c, y = total_bike_count)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm") +
  labs(x = "Temperature (C)",
       y = "Total Bike Rentals",
       title = "Temperature vs Bike Rentals")
```

As with the sunnier weather, temperature seems directly proportional to bike rentals.

```{r Corr Plot}
num_vars <- daily_bike_data[ , 4:12]

corrplot(cor(num_vars))
```

By looking at the correlation plot above, the total bike rental count has a strong, positive correlation with the average temperature and solar radiation. This makes sense as more bikes being rented on a warm, sunny day rather than a cold, cloudy day.

#### Split the Data

Splitting `daily_bike_data` into training and testing sets (75/25 split).

```{r Splitting}
set.seed(75)

bike_split <- initial_split(daily_bike_data, prop = 0.75, strata = seasons)
bike_train <- training(bike_split)
bike_test <- testing(bike_split)
```

Create a 10 fold cross-validation (CV) split on the training set.

```{r 10 Fold CV}
bike_folds <- vfold_cv(bike_train, v = 10)
```

#### Fitting MLR Models

Bike data is ready for fitting linear regression models! First, create some recipes.

For the 1st recipe, ignore the `date` variable for modeling and instead use it to create a weekday/weekend (factor) variable.

```{r Recipe 1}
bike_recipe_1 <- recipe(total_bike_count ~ ., data = bike_train) |>
  update_role(date, new_role = "ID") |> # Give 'date' an ID to keep it, but not for modeling.
  step_date(date, features = "dow", label = TRUE) |> # Pull out week days from date.
  
  # Create new weekday/weekend factor variable.
  step_mutate(
    day_type = factor(if_else(date_dow %in% c("Sat", "Sun"), "weekend", "weekday"))) |>
  
  step_rm(date_dow) |> # Remove intermediate day of week variable.
  step_normalize(all_numeric_predictors()) |> # Standardize num vars.
  step_dummy(all_nominal_predictors()) |> # Dummy variables for cat vars.
  step_zv(all_predictors()) # For zero-variance predictor that has no information within the column.

# Check summary
summary(bike_recipe_1)
```

For the 2nd recipe, repeat the steps above, as in recipe 1, but now add in **interactions** between seasons and holiday, seasons and temp, temp and rainfall.

```{r Recipe 2}
bike_recipe_2 <- recipe(total_bike_count ~ ., data = bike_train) |>
  update_role(date, new_role = "ID") |> # Give 'date' an ID
  step_date(date, features = "dow", label = TRUE) |> # Pull out week days
  
  # Create weekday/weekend variable
  step_mutate(
    day_type = factor(if_else(date_dow %in% c("Sat", "Sun"), "weekend", "weekday"))) |>
  
  step_rm(date_dow) |> # Remove intermediate variable
  
    step_normalize(all_numeric_predictors()) |> # Standardize
  
  step_dummy(all_nominal_predictors()) |> # Dummy variables
  
  # Interactions
  step_interact( ~ starts_with("seasons_"):starts_with("holiday_")) |> # seasons x holiday
  step_interact( ~ starts_with("seasons_"):starts_with("avg_temp_c")) |> # seasons x temp
  step_interact( ~ starts_with("avg_temp_c"):starts_with("total_rainfall_mm")) |> # temp x rainfall
  
  step_zv(all_predictors())

summary(bike_recipe_2)
```

For the 3rd and final recipe, repeat the steps above, as in recipe 2, but now add in **quadratic terms** for each numeric predictor.

```{r Recipe 3}
bike_recipe_3 <- recipe(total_bike_count ~ ., data = bike_train) |>
  update_role(date, new_role = "ID") |> # Give 'date' an ID
  step_date(date, features = "dow", label = TRUE) |> # Pull out week days
  
  # Create weekday/weekend variable
  step_mutate(
    day_type = factor(if_else(date_dow %in% c("Sat", "Sun"), "weekend", "weekday"))) |>
  
  step_rm(date_dow) |> # Remove intermediate variable
  
  step_poly(all_numeric_predictors(), degree = 2) |> # ADD IN QUADRATIC TERMS
  
    step_normalize(all_numeric_predictors()) |> # Standardize

    step_dummy(all_nominal_predictors()) |> # Dummy variables

  
# Interactions
  step_interact( ~ starts_with("seasons_"):starts_with("holiday_")) |> # seasons x holiday
  step_interact( ~ starts_with("seasons_"):starts_with("avg_temp_c")) |> # seasons x temp
  step_interact( ~ starts_with("avg_temp_c"):starts_with("total_rainfall_mm")) |> # temp x rainfall
  
  step_zv(all_predictors())

summary(bike_recipe_3)
```

Finally, set up linear model fit to use the “lm” engine. Fit the models using 10 fold CV, then consider the training set CV error to choose a best model.

```{r Additional Packages}
library(parsnip)
library(workflows)
library(broom)
```

Here, start by building a model specification. As stated earlier, using a *linear model* on bike data.

```{r Model Specification}
lm_mod <- 
  linear_reg() |> 
  set_engine("lm")
```

Then, create some model workflows, which pair models and recipes together. Do this for all three recipes.

```{r Workflows}
bike_wf1 <- 
  workflow() |>
  add_model(lm_mod) |>
  add_recipe(bike_recipe_1)

bike_wf2 <- 
  workflow() |>
  add_model(lm_mod) |>
  add_recipe(bike_recipe_2)

bike_wf3 <- 
  workflow() |>
  add_model(lm_mod) |>
  add_recipe(bike_recipe_3)
```

Next, fit the models using 10 fold CV.

```{r}
set.seed(75)

cv_wf1 <- bike_wf1 |>
  fit_resamples(bike_folds)

cv_wf2 <- bike_wf2 |>
  fit_resamples(bike_folds)

cv_wf3 <- bike_wf3 |>
  fit_resamples(bike_folds)
```

Now, look at the training set CV error to choose the best model.

```{r Observe RMSE}
collect_metrics(cv_wf1)
collect_metrics(cv_wf2)
collect_metrics(cv_wf3) # Lowest RMSE
```

Since the third model seems to have the lowest RMSE, fit this model to the entire training data set. Then, compute the RMSE metric on the test set.

```{r}
best_fit <- bike_wf3 |>
  last_fit(bike_split)

best_fit |> collect_metrics()
```

To end, obtain the final model (fit on the entire training set) coefficient table.

```{r}
final_model_coeffs <- best_fit |>
  extract_fit_parsnip() |>
  tidy()

final_model_coeffs
```

## HW 8 Begins

For the overall best model, fit it to the entire data set!

```{r}
library(glmnet)
library(baguette)
library(ranger)
library(rpart.plot)
```


In this section, this is continuation with the previous code and analysis done in homework 7. Using the first recipe that was written, several new models are added. Adding a **LASSO model** to fit on our bike data. In this case, use linear regression, but using the "glmnet" engine instead of "lm".

```{r LASSO spec}
LASSO_recipe <- bike_recipe_1

LASSO_spec <- linear_reg(penalty = tune(), mixture = 1) |>
  set_engine("glmnet")
```

Next, create the workflow.
```{r LASSO wkf}

LASSO_wkf <- workflow() |>
  add_recipe(LASSO_recipe) |>
  add_model(LASSO_spec)
LASSO_wkf
```

Below, use the `tune_grid()` function. This function allows to fit the model to CV folds but specify the set of tuning parameters to consider.
```{r LASSO grid}
set.seed(75)

LASSO_grid <- LASSO_wkf |>
  tune_grid(resamples = bike_folds,
            grid = grid_regular(penalty(), levels = 200)) 
```

Now, these metrics are computed across the folds for each the 200 values of the tuning parameter.
```{r LASSO metric}
LASSO_grid |>
  collect_metrics() |>
  filter(.metric == "rmse")
```

Find the best model from the LASSO models and fit it to the entire training data set and see how it predicts on the test set.

```{r LASSO finalize & fit}
best_LASSO <- LASSO_grid |> 
  select_best(metric = "rmse")

LASSO_final <- LASSO_wkf |> 
  finalize_workflow(best_LASSO) |>
  last_fit(bike_split, metrics = metric_set(rmse, mae))
```


## Regression Tree Model

Repeat the same steps, but instead use a **regression tree model**. First, define the recipe and model specifications.
```{r Tree spec}

tree_recipe <- bike_recipe_1

tree_mod <- decision_tree(tree_depth = tune(),
                          min_n = 20,
                          cost_complexity = tune()) |>
  set_engine("rpart") |>
  set_mode("regression")
```

Next, create the workflow.

```{r Tree wkf}
tree_wkf <- workflow() |>
  add_recipe(tree_recipe) |>
  add_model(tree_mod)
tree_wkf
```

Then, fit the workflow to CV folds.
```{r Tree fit wkf to folds}
set.seed(75)

tree_grid <- grid_regular(cost_complexity(),
                          tree_depth(),
                          levels = c(10, 5))

tree_fits <- tree_wkf |> 
  tune_grid(resamples = bike_folds,
            grid = tree_grid)
tree_fits
```

Now that the model has been tuned, collect metrics to see how the models performed.

```{r Tree metrics}
tree_fits |>
  collect_metrics() |>
  filter(.metric == "rmse") |>
  arrange(mean)

best_tree <- tree_fits |> 
  select_best(metric = "rmse")
best_tree
```

Finally, finalize the workflow and fit it to the entire training data set.

```{r Tree finalize & fit}
tree_final <- tree_wkf |>
  finalize_workflow(best_tree) |>
  last_fit(bike_split, metrics = metric_set(rmse, mae))

tree_final |>
  collect_metrics()
```


## Bagged Tree Model

Repeat the same steps again, but instead use a **bagged tree model**. First, define the recipe, model specifications, and create the workflow.
```{r Bag spec & wkf}
bag_recipe <- bike_recipe_1

# Model specification
bag_spec <- bag_tree(tree_depth = 5, min_n = 10, cost_complexity = tune()) |>
  set_engine("rpart") |>
  set_mode("regression")

# Workflow
bag_wkf <- workflow() |>
  add_recipe(bag_recipe) |>
  add_model(bag_spec)
bag_wkf
```

Next, fit the workflow to the CV folds and check the metrics across the folds.
```{r Bag fit to folds & check metrics}
set.seed(75)

bag_grid <- grid_regular(cost_complexity(), levels = 15)

bag_fit <- bag_wkf |>
  tune_grid(resamples = bike_folds,
            grid = bag_grid)

# Collect metrics
bag_fit |>
  collect_metrics() |>
  filter(.metric == "rmse") |>
  arrange(mean)

# Check metric performance
best_bag <- bag_fit |> 
  select_best(metric = "rmse")
best_bag
```

Finally, finalize and fit the model onto the training data set.
```{r Bag finalize & fit}
bag_final <- bag_wkf |>
  finalize_workflow(best_bag) |>
  last_fit(bike_split, metrics = metric_set(rmse, mae))

bag_final |>
  collect_metrics()
```


## Random Forest Model

Repeat the same steps again, but instead use a **random forest model**. First, define the recipe, model specifications, and create the workflow.
```{r RF spec & wkf}
rf_recipe <- bike_recipe_1

# Model specification
rf_spec <- rand_forest(mtry = tune()) |>
  set_engine("ranger", importance = "impurity") |> # Added level for importance
  set_mode("regression")

# Workflow
rf_wkf <- workflow() |>
  add_recipe(rf_recipe) |>
  add_model(rf_spec)
rf_wkf
```

Next, fit the workflow to the CV folds and checking the metrics across the folds.
```{r RF fit to folds & check metrics}
set.seed(75)

rf_fit <- rf_wkf |>
  tune_grid(resamples = bike_folds,
            grid = 7)

# Collect metrics
rf_fit |>
  collect_metrics() |>
  filter(.metric == "rmse") |>
  arrange(mean)

# Check metric performance
best_rf <- rf_fit |> 
  select_best(metric = "rmse")
best_rf
```

Finally, finalize and fit the model onto the training data set.
```{r RF finalize & fit}
rf_final <- rf_wkf |>
  finalize_workflow(best_rf) |>
  last_fit(bike_split, metrics = metric_set(rmse, mae))

rf_final |>
  collect_metrics()
```


## Comparing All Final Models

In this last section, compare all final models on the test set using both RMSE and MAE (mean absolute error).
```{r}
# Re-fit MLR with both metrics included
mlr_final <- bike_wf3 |>
  last_fit(bike_split, metrics = metric_set(rmse, mae))

all_metrics <- rbind(
  mlr_final |>
    collect_metrics() |>
    mutate(Model = "MLR", .before = ".metric"),
  LASSO_final |>
    collect_metrics() |>
    mutate(Model = "LASSO", .before = ".metric"),
  tree_final |>
    collect_metrics() |>
    mutate(Model = "REGRESSION_TREE", .before = ".metric"),
  bag_final |>
    collect_metrics() |>
    mutate(Model = "BAGGED_TREE", .before = ".metric"),
  rf_final |>
    collect_metrics() |>
    mutate(Model = "RANDOM_FOREST", .before = ".metric")
)

all_metrics |>
  filter(.metric %in% c("rmse", "mae")) |> 
  arrange(.estimate)
```

According to RMSE, we can see that the best model for our data is the Random Forest Model!

## Final Model Fits & Summaries

For the LASSO and MLR models, report the final coefficient tables.
```{r coefficient tables}
tidy(mlr_final |> extract_fit_parsnip()) |> arrange(desc(abs(estimate))) # MLR
tidy(LASSO_final |> extract_fit_parsnip()) |> arrange(desc(abs(estimate))) # LASSO
```

For the regression tree model, create a plot of the final fit.
```{r Tree final fit plot}
tree_obj <- tree_final |> 
   extract_workflow() |> 
   extract_fit_engine()
 
 # Plot
 rpart.plot(tree_obj, 
            roundint = FALSE,
            main = "Regression Tree Final Fit")
```

For the bagged tree and random forest models, produce variable importance plots.
```{r Bagged variable importance}
# Extract final model fit
bag_final_model <- bag_final |> extract_workflow() |> extract_fit_engine()

bag_final_model$imp |>
  mutate(term = factor(term, levels = term)) |> 
  ggplot(aes(x = term, y = value)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Bagged Tree Variable Importance",
       x = "Variables",
       y = "Importance")
```

```{r RF variable importance}
# Extract final model fit
rf_final_model <- rf_final |> extract_workflow() |> extract_fit_engine()

rf_import <- rf_final_model$variable.importance |> 
  sort(decreasing = TRUE)

# Tibble for plotting
rf_import_df <- tibble(
  variable = names(rf_import),
  importance = rf_import
)

rf_import_df |> 
  ggplot(aes(x = reorder(variable, importance), y = importance)) +
  geom_col(fill = "forestgreen") +
  coord_flip() +
  labs(title = "Random Forest Variable Importance",
       x = "Variables",
       y = "Importance")
```
In both the bagged tree and random forest variable importance plots above, the most important predictor for the number of bike rentals is the average temperature (c) outside.

Fit the overall best model to the entire data set. This will be the Random Forest Model.
```{r}
final_rf_wkf <- rf_wkf |>
  finalize_workflow(best_rf)

final_rf_fit <- final_rf_wkf |>
  fit(data = daily_bike_data)

final_rf_model <- final_rf_fit |>
  extract_fit_engine()

final_rf_model
```
The final random forest model is a regression model that was trained on the full bike data set using 500 trees. This model is able to predict daily bike rental counts pretty accurately, with about 92.2% of the variance in daily bike rentals. It also achieved an Out-of-Bag (OOB) Root Mean Squared Error (RMSE) of roughly 2,767 bikes, which is highly consistent with the error observed on our test set. As seen previously in the variable importance plot for this model, the average temperature (c) seems to be the most important variable in the data set when predicting daily bike rental counts.
