---
title: "Homework 8"
format: html
---


```{r}
library(tidyverse)
library(tidymodels)
library(ggplot2)
```


```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(corrplot)
library(tidymodels)
library(glmnet)
library(tree)
library(rpart.plot)
library(baguette)
library(ranger)
```

## Read Data

```{r}
bikeDf = read.csv("SeoulBikeData.csv", skip = 1, col.names = c("Date",
                                                               "RentedBikeCount",
                                                               "Hour",
                                                               "Temperature",
                                                               "Humidity",
                                                               "WindSpeed",
                                                               "Visibility",
                                                               "DewPoint",
                                                               "SolarRadiation",
                                                               "Rainfall",
                                                               "Snowfall",
                                                               "Seasons",
                                                               "Holiday",
                                                               "FunctioningDay"))

head(bikeDf)
```

## EDA

### Check Variable Types and Cast Proper Variable Types

```{r}
print(str(bikeDf))
```
"Holiday", "Seasons", and "FunctioningHoliday" should be factors. "Date" should be a DateTime.

```{r}
bikeDf = as.tibble(bikeDf) |> mutate(Holiday = as.factor(Holiday),
                                     Seasons = as.factor(Seasons),
                                     FunctioningDay = as.factor(FunctioningDay),
                                     Date = as.Date(Date))

head(bikeDf)
```

### Assess Missingness

```{r}

sum_na <- function(column){
  sum(is.na(column))
}
bikeDf_na <- bikeDf |>
  summarize(across(everything(), sum_na))

bikeDf_na

```
No data is missing! Lovely

Numeric Summary Statistics:

```{r}
numVarSummaries = bikeDf |> summarize(across(where(is.numeric), 
                                      list("mean" = mean, "standardDeviation" = sd),
                                      .names = "{.fn}_{.col}"))

numVarSummaries
  
```
### Scatter Plot between numeric variables

```{r}
corrplot(cor(bikeDf |> select((where(is.numeric)))))
```
### Box Graphs for Factors

```{r}

ggplot(data = bikeDf |> select(Holiday) |>
         group_by(Holiday), aes(x = Holiday)) + geom_bar()

ggplot(data = bikeDf |> select(Seasons) |>
         group_by(Seasons), aes(x = Seasons)) + geom_bar()

ggplot(data = bikeDf |> select(FunctioningDay) |>
         group_by(FunctioningDay), aes(x = FunctioningDay)) + geom_bar()
```

FuntioningDay and Holiday Variables are skewed towards "No" and "No Holiday"

### Scatter Plots:

```{r}

g = ggplot(data = bikeDf, aes(y = RentedBikeCount))

g + geom_point(aes(x = Temperature,color = Seasons))

g + geom_point(aes(x = Temperature,color = FunctioningDay))

```
It looks like when there is not a  Functioning Day, the bike count drops to 0. Let's subset these out. 

### Generate Summary Data:

```{r}
bikeDfSummary = bikeDf |>
  filter(FunctioningDay == "Yes") |>
  group_by(Date, Seasons, Holiday) |>
  summarize(sumBikeCount = sum(RentedBikeCount),
            sumRainfall = sum(Rainfall),
            sumSnowfall = sum(Snowfall),
            meanTemperature = mean(Temperature),
            meanHumidity = mean(Humidity),
            meanWindSpeed = mean(WindSpeed),
            meanVisibility = mean(Visibility),
            meanDewPoint = mean(DewPoint),
            meanSolarRadiation = mean(SolarRadiation),
            )

bikeDfSummary
  
```

```{r}
head(bikeDfSummary)
```


### Summary Stats of New Data:

```{r}
numVarSummaries = bikeDfSummary |> ungroup() |> summarize(across(where(is.numeric), 
                                      list("mean" = mean, "standardDeviation" = sd),
                                      .names = "{.fn}_{.col}"))

numVarSummaries
```


### A few scatter looking at predictor variables vs bike count:

```{r}
g = ggplot(data = bikeDfSummary, aes(y = sumBikeCount))

g + geom_point(aes(x = meanTemperature,color = Seasons))

g + geom_point(aes(x = meanTemperature,color = Holiday))

g + geom_point(aes(x = meanTemperature,color = Holiday))
```
Certainly the least amount of bikes being rented during the cold weather (winter) and the most being in warm weather (to a degree, until it's too hot). Relationship between temp and bike count appears similar during fall and spring.

### Correlation between numeric variables

```{r}
numVars = bikeDfSummary |> ungroup() |> select((where(is.numeric)))

print(cor(numVars))

corrplot(cor(numVars))

```
Highest positive correlated variables with bike count are
- meanTemperature
- mewsDewPoint (Also highly correlated with meanTemperature)
- meanSolar Radaition

It seems like people like riding bikes during warm and sunny days.


## Modeling

### Create Train / Test Split
```{r}
head(bikeDfSummary)
```


```{r}

bikeDfSummary = bikeDfSummary |> ungroup()

bikeSplit = initial_split(bikeDfSummary, prop = .75, strata = Seasons)
bikeTrain = training(bikeSplit)
bikeTest = testing(bikeSplit)
```


### Create CV Splits

```{r}
bikeCVFolds = vfold_cv(bikeTrain, 10)
```


### First Recipe for Cleaning Data


```{r}



recipe1 = recipe(data = bikeTrain, sumBikeCount ~.) |> 
  step_mutate(isWeekend = factor(ifelse(weekdays(Date) %in% c("Saturday", "Sunday"), "N", "Y"))) |>
  update_role(Date, new_role = "ID") |>
  step_dummy(Seasons, isWeekend, Holiday) |>
  step_normalize(all_numeric())
  
```



### Add Interfactions to a new recipe

```{r}
recipe2 = recipe(data = bikeTrain, sumBikeCount ~.) |> 
  step_mutate(isWeekend = factor(ifelse(weekdays(Date) %in% c("Saturday", "Sunday"), "N", "Y"))) |>
  update_role(Date, new_role = "ID") |>
  step_dummy(Seasons, isWeekend, Holiday) |>
  step_normalize(all_numeric()) |>
  step_interact(terms = ~ starts_with("Seasons"):starts_with("Holiday")) |>
  step_interact(terms = ~ starts_with("Seasons"):meanTemperature) |>
  step_interact(terms = ~ sumRainfall:meanTemperature)

```


### Add Quadratic Terms to Numeric Predictors

```{r}
recipe3 = recipe(data = bikeTrain, sumBikeCount ~.) |> 
  step_mutate(isWeekend = factor(ifelse(weekdays(Date) %in% c("Saturday", "Sunday"), "N", "Y"))) |>
  update_role(Date, new_role = "ID") |>
  step_dummy(Seasons, isWeekend, Holiday) |>
  step_normalize(all_numeric(), -all_outcomes()) |>
  step_interact(terms = ~ starts_with("Seasons"):starts_with("Holiday")) |>
  step_interact(terms = ~ starts_with("Seasons"):meanTemperature) |>
  step_interact(terms = ~ sumRainfall:meanTemperature) |>
  step_poly(all_of(c("meanTemperature", "meanHumidity", "meanWindSpeed",
                    "meanDewPoint", "meanVisibility", "meanSolarRadiation",
                    "sumRainfall", "sumSnowfall")),
          degree = 2)
  
  


```

### Create MLR Model

```{r}
mlrModel <- linear_reg() |> set_engine("lm")
```

### Fit all 3 Models and Evaluate with 10 Fold CV

```{r}

bikeFitWorkflow1 = workflow() |>
  add_recipe(recipe1) |>
  add_model(mlrModel) |>
  fit_resamples(bikeCVFolds)

results1 = bikeFitWorkflow1 |> collect_metrics()

results1

```


# Beginning of HW 8

Objective:

10 Fold CV for the following Models:
-    LASSO + Coef Table
-    Regression Tree + Plot of Final Tree
-    Bagged Tree + Variable Importance Plot
-    Random Forest + Variable Importance Plot

## MLR Model

```{r}

mlrFitWorkflow = workflow() |>
  add_recipe(recipe1) |>
  add_model(mlrModel)


MLR_final <-  mlrFitWorkflow |>
  fit(bikeTrain)
tidy(MLR_final)



```


## LASSO

```{r}
lassoModel <- linear_reg(penalty = tune(), mixture = 1) |>
  set_engine("glmnet")

lassoWorkflow <- workflow() |>
  add_recipe(recipe1) |>
  add_model(lassoModel)

#A warning will occur for one value of the tuning parameter, safe to ignore
LASSO_grid <- lassoWorkflow |>
  tune_grid(resamples = bikeCVFolds,
            grid = grid_regular(penalty(), levels = 200),
            save_workflow = TRUE) 
LASSO_grid


LASSO_grid |>
  collect_metrics() |>
  filter(.metric == "rmse") |>
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_line()

```
```{r}

lowest_rmse <- LASSO_grid |>
  select_best(metric = "rmse")
lowest_rmse

lassoWorkflow |>
  finalize_workflow(lowest_rmse)

#fit it to the entire training set to see the model fit
LASSO_final <- lassoWorkflow |>
  finalize_workflow(lowest_rmse) |>
  fit(bikeTrain)
tidy(LASSO_final)


```

## Regression Tree

```{r}

## Model Spec
treeModel <- decision_tree(tree_depth = tune(),
                          min_n = 20,
                          cost_complexity = tune()) |>
  set_engine("rpart") |>
  set_mode("regression")

## Workflow
tree_wkf <- workflow() |>
  add_recipe(recipe1) |>
  add_model(treeModel)

tree_grid <- grid_regular(cost_complexity(),
                          tree_depth(),
                          levels = c(10, 5))

tree_fits <- tree_wkf |> 
  tune_grid(resamples = bikeCVFolds,
            grid = tree_grid)

tree_fits |>
  collect_metrics()


```

### Extract Best Fit:

```{r}
tree_best_params <- select_best(tree_fits, metric = "rmse")
tree_best_params
```
### Obtain Best Model Fit on Training Set
```{r}
tree_final_wkf <- tree_wkf |>
  finalize_workflow(tree_best_params)

tree_final_fit <- tree_final_wkf |>
  last_fit(bikeSplit)
tree_final_fit

tree_final_model <- extract_workflow(tree_final_fit) 
tree_final_model

tree_final_model |>
  extract_fit_engine() |>
  rpart.plot::rpart.plot(roundint = FALSE)
```

## Bagged Tree

```{r}

## Model

baggedTreeModel <- bag_tree(tree_depth = 5, min_n = 10, cost_complexity = tune()) |>
  set_engine("rpart") |>
  set_mode("regression")

## Workflow

baggedTreeWorkflow = workflow() |>
  add_recipe(recipe1) |>
  add_model(baggedTreeModel)

## Fit object with Cross Validaiton

bag_fit <- baggedTreeWorkflow |>
  tune_grid(resamples = bikeCVFolds, 
            grid = grid_regular(cost_complexity(),
                                levels = 15),
            metrics = metric_set(rmse, mae))
bag_fit

## Obtain Best Fit

bag_best_params <- select_best(bag_fit, metric = "rmse")
bag_best_params

## Fit Model on Best Fit

bag_final_wkf <- baggedTreeWorkflow |>
  finalize_workflow(bag_best_params)
bag_final_fit <- bag_final_wkf |>
  last_fit(bikeSplit, metrics = metric_set(rmse, mae))

## Variable Importance Plot

bag_final_model <- extract_fit_engine(bag_final_fit) 
bag_final_model$imp |>
  mutate(term = factor(term, levels = term)) |>
  ggplot(aes(x = term, y = value)) + 
  geom_bar(stat ="identity") +
  coord_flip()
```

## Random Forest

```{r}

## Model Def

rfModel <- rand_forest(mtry = tune()) |>
  set_engine("ranger") |>
  set_mode("regression")

## Workflow

rfWorkflow = workflow() |>
  add_recipe(recipe1) |>
  add_model(rfModel)

## Fit on CV Object

rf_fit <- rfWorkflow|>
  tune_grid(resamples = bikeCVFolds, 
            grid = 7,
            metrics = metric_set(rmse, mae))

## Obtain Best Paramters
rf_best_params <- select_best(rf_fit, metric = "rmse")
rf_best_params


## Fit Test Set on Best Model
rf_final_wkf <- rfWorkflow |>
  finalize_workflow(rf_best_params)
rf_final_fit <- bag_final_wkf |>
  last_fit(bikeSplit, metrics = metric_set(rmse, mae))

## Variable Importance Plot

cells_res <-
    rfWorkflow %>%
    tune_grid(resamples = bikeCVFolds, 
            grid = 7,
            metrics = metric_set(rmse, mae))


```


## Compare All Models

```{r}
MLR_metrics = bikeFitWorkflow1 |> fit(bikeTest)

MLR_metrics
LASSO_metrics
```

