---
title: "Homework 8"
author: Bryan Sandor, Merilyn Thomas, and Tyler White
format: html
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
---

# Previous Document (Homework 7)
## Preamble
The purpose of this homework is to practice with fitting linear models in `R`. The data is from the UCI Machine Learning Repository concerning bike sharing rentals, available [here](https://www4.stat.ncsu.edu/online/datasets/SeoulBikeData.csv). First we "summon" the necessary packages.

```{r packages}
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(parsnip)
library(workflows)
library(broom)
```


## Reading Data
We will read in the data directly from the source, rather than downloading it.

```{r read_data}
# an error results in this command, "fixed" by the call to "locale" command
raw_bike_data <- read_csv(
  "https://www4.stat.ncsu.edu/online/datasets/SeoulBikeData.csv",
  locale = locale(encoding = "latin1")
  )
```

## EDA
### Tidying
Before we start analysis, we will need to clean up the data.

1. We begin by checking the data for missing values. The below command sums across all columns the counts of `NA` values.

```{r missingness}
colSums(is.na(raw_bike_data))
```

We observe there are no explicit `NA` values within the data.

2. Now we check the data to ensure each of the variable values "make sense" in the context of the type.

```{r column_types}
summary(raw_bike_data)
```

Both numeric and categorical data appear to be in appropriate ranges.

3. The date column was incorrectly read in as a `char` type so we will convert it to the appropriate `date` type. We use the `dmy` (for "day-month-year") command from the `lubridate` package.

```{r convert_date_column}
clean_bike_data <- raw_bike_data

clean_bike_data$Date <- dmy(clean_bike_data$Date)

head(clean_bike_data$Date)
```

4. Now we adjust the remaining `char` type variables to factors.

```{r change_to_factors}
clean_bike_data$Seasons <- as.factor(clean_bike_data$Seasons)
clean_bike_data$Holiday <- as.factor(clean_bike_data$Holiday)
clean_bike_data$`Functioning Day` <- as.factor(clean_bike_data$`Functioning Day`)

clean_bike_data |> select(c(Seasons, Holiday, `Functioning Day`))
```

5. Finally, we will rename all the variables to more "friendly" names.

```{r renaming}
clean_bike_data <- clean_bike_data |> rename(
  date = Date,
  rent_count = `Rented Bike Count`,
  hour = Hour,
  temp = `Temperature(°C)`,
  humidity = `Humidity(%)`,
  wind_speed = `Wind speed (m/s)`,
  visibility = `Visibility (10m)`,
  dew_point = `Dew point temperature(°C)`,
  sun_radiation = `Solar Radiation (MJ/m2)`,
  rainfall = `Rainfall(mm)`,
  snowfall = `Snowfall (cm)`,
  season = Seasons,
  holiday = Holiday,
  functional_hour = `Functioning Day`
)
```

### Fine-Tuning
Now that the data is formatted ideally, we can do some basic analysis.

6. First, let's observe summary statistics.

```{r summary_stats}
summary(clean_bike_data)
```

There are `295` observations occurring within a non-functional hour, rendering those observations irrelevant to our analysis, so we will need to remove them from our data set.

```{r remove_bad_observations}
functional_bike_data <- clean_bike_data |> filter(
  functional_hour == "Yes"
)

summary(functional_bike_data$functional_hour)
```

7. Now that we have a fully functional, tidied data set, let's perform some analyses on it. We will summarize across the hours so each day has one observation.

```{r target_data_set}
bike_data <- functional_bike_data |> group_by(date, season, holiday) |>
  summarize(
    total_rent_count = sum(rent_count),
    total_rainfall = sum(rainfall),
    total_snowfall = sum(snowfall),
    mean_temp = mean(temp),
    mean_humidity = mean(humidity),
    mean_wind_speed = mean(wind_speed),
    mean_visibility = mean(visibility),
    mean_dew_point = mean(dew_point),
    mean_sun_radiation = mean(sun_radiation)
  )

head(bike_data)
```

### Analysis
#### Numeric Summary
8. We have the data set we need, so let's perform some basic analyses on them, beginning with a simple summary.

```{r final_summary}
summary(bike_data)
```

#### Graphs
Now let's examine some plots. First, let's relate rainfall to total rentals.

```{r rainfall_v_rentals_graph}
ggplot(data = bike_data,
       aes(x = total_rainfall, y = total_rent_count)) +
  geom_jitter(width = 1) +
  geom_smooth(method = lm) +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 40000)) +
  labs(x = "Total Rainfall (mm)",
       y = "Total Bike Rentals",
       title = "Rainfall vs Bike Rentals")
```

It looks like an inverse relation exists between rainfall and rentals. Now let's do the same but for snowfall. One would assume a similar relationship exists.

```{r snowfall_v_rentals_graph}
ggplot(data = bike_data,
       aes(x = total_snowfall, y = total_rent_count)) +
  geom_jitter(width = 1) +
  geom_smooth(method = lm) +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 40000)) +
  labs(x = "Total Snowfall (cm)",
       y = "Total Bike Rentals",
       title = "Snowfall vs Bike Rentals")
```

Not surprisingly, this seems more stark; after all, snow is more dangerous for bike rides and some people even prefer to ride in the rain. Let's examine the rentals across each of the four seasons.

```{r seasonal_rentals_graph}
ggplot(data = bike_data,
       aes(x = season, y = total_rent_count)) +
  geom_col() +
  labs(x = "Season",
       y = "Total Bike Rentals",
       title = "Seasonal Bike Rentals")
```

It appears people prefer sunny weather or cooler, but not frigid, temperatures, and this is practical. Let's touch on this last thought with two more comparisons: solar radiation and temperature, respectively.

```{r solar_radiation_vs_rentals_graph}
ggplot(data = bike_data,
       aes(x = mean_sun_radiation, y = total_rent_count)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(x = "Solar Radiation (MJ/m^2)",
       y = "Total Bike Rentals",
       title = "Solar Radiation vs Bike Rentals")
```

Indeed, sunnier weather seems directly proportional to bike rentals, albeit not within as narrow a band as our rain or snow fall.

```{r temperature_vs_rentals_graph}
ggplot(data = bike_data,
       aes(x = mean_temp, y = total_rent_count)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(x = "Temperature (C)",
       y = "Total Bike Rentals",
       title = "Temperature vs Bike Rentals")
```

As with the sunnier weather, temperature seems directly proportional to bike rentals.

#### Correlation
Finally, let's explore combinations of correlations among numeric variables.

```{r correlation}
cor(bike_data[,4:12]) # uses just the numeric columns
```

The highest correlations with the number of rentals appears to be temperature, solar radiation, and a variable we did not explore in the graphs: dew point, with the weakest as humidity, visibility, and wind speed. It surprises the author that rainfall and snowfall were not more highly correlated.

## Modelling
### Split the Data
First we split the data into a training and test set (75/25 split), stratified across the seasons, creating a 10 fold cross-validation (CV) split.

```{r data_split}
set.seed(44)

bike_split <- initial_split(bike_data,
                            prop = 0.75,
                            strata = season)

train_data <- training(bike_split)
test_data <- testing(bike_split)

cv_folds <- vfold_cv(train_data,
                     v = 10,
                     strata = season)
```

### Fitting MLR Models
#### First Recipe
For our first recipe, we want to remove the `date` variable as a predictor, but use it to create a new variable designating whether the date fell on a weekday or weekend. Then, we standardize all the numeric data (converts all the data to a $z$-score for its corresponding variable) and craft a dummy variable for the season, holiday, and our new workweek factors.

```{r first_recipe}
bike_rec_1 <- recipe(total_rent_count ~ .,
                     data = train_data) |>
  update_role(date,
              new_role = "date") |> # removes date as predictor
  step_date(date,
            features = "dow",
            label = TRUE) |> # returns the day of the week (dow)
  step_mutate(workweek = factor(
    if_else(date_dow %in% c("Sat", "Sun"),
    "weekend",
    "weekday"))) |> # creates new variable as factor for weekend or weekday
  step_rm(date_dow) |> # removes temporary variable from previous step
  step_normalize(all_numeric_predictors()) |> # standardizes all numeric data
  step_dummy(all_nominal_predictors()) # "dummifies" all categorical data

summary(bike_rec_1) # indicates roles of variables
```

#### Second Recipe
The second recipe is identical to the first, except we add in interactions between seasons and holiday; seasons and temp; and temp and rainfall.

```{r second_recipe}
bike_rec_2 <- recipe(total_rent_count ~ .,
                     data = train_data) |>
  update_role(date,
              new_role = "date") |>
  step_date(date,
            features = "dow",
            label = TRUE) |>
  step_mutate(workweek = factor(
    if_else(date_dow %in% c("Sat", "Sun"),
    "weekend",
    "weekday"))) |>
  step_rm(date_dow) |>
  step_normalize(all_numeric_predictors()) |>
  step_dummy(all_nominal_predictors()) |> # end of first recipe
  step_interact( ~ starts_with("season"):starts_with("holiday")) |>
  step_interact( ~ starts_with("season"):starts_with("mean_temp")) |>
  step_interact( ~ starts_with("mean_temp"):starts_with("total_rainfall"))
```

#### Third Recipe
The last recipe is identical to the second, except we now include quadratic terms for each of the 10 numeric predictors.

```{r third_recipe}
bike_rec_3 <- recipe(total_rent_count ~ .,
                     data = train_data) |>
  update_role(date,
              new_role = "date") |>
  step_date(date,
            features = "dow",
            label = TRUE) |>
  step_mutate(workweek = factor(
    if_else(date_dow %in% c("Sat", "Sun"),
    "weekend",
    "weekday"))) |>
  step_rm(date_dow) |>
  step_poly(all_numeric_predictors(), degree = 2) |> # adds quadratic terms
  step_normalize(all_numeric_predictors()) |>
  step_dummy(all_nominal_predictors()) |> 
  step_interact( ~ starts_with("season"):starts_with("holiday")) |>
  step_interact( ~ starts_with("season"):starts_with("mean_temp")) |>
  step_interact( ~ starts_with("mean_temp"):starts_with("total_rainfall"))
```

#### Model Fit
Now we can set up our linear model fit to use the "lm" engine.

```{r lm_fit}
bike_lm <- linear_reg() |>
  set_engine("lm")
```

We continue by making workflows for each of the three recipes.

```{r workflows}
bike_workflow_1 <- workflow() |>
  add_model(bike_lm) |>
  add_recipe(bike_rec_1)
  
bike_workflow_2 <- workflow() |>
  add_model(bike_lm) |> 
  add_recipe(bike_rec_2)
  
bike_workflow_3 <- workflow() |>
  add_model(bike_lm) |> 
  add_recipe(bike_rec_3)
```

Then we fit the models using our previous 10 fold CV.

```{r fit_models}
cv_bike_1 <- bike_workflow_1 |>
  fit_resamples(cv_folds)

cv_bike_2 <- bike_workflow_2 |>
  fit_resamples(cv_folds)

cv_bike_3 <- bike_workflow_3 |>
  fit_resamples(cv_folds)
```

We need to see which of the three models works best by examining the CV errors.

```{r CV_errors}
collect_metrics(cv_bike_1)
collect_metrics(cv_bike_2)
collect_metrics(cv_bike_3)
```

## Final Model
We see the third model has the lowest root-mean squared error (RMSE) and thus we will use it as a model to fit on the entire training set.

```{r best_model_fit}
bike_best_model <- bike_workflow_3 |>
  last_fit(bike_split)

collect_metrics(bike_best_model)
```

Finally, to "visualize" the model, we will extract the coefficients.

```{r model_coefficients}
bike_best_model |> extract_fit_parsnip() |>
  tidy() |>
  print(n = Inf)
```

# Homework 8 (Current Document)